# Methods

Workflows for the project have been classified into planning, fish passage assessments, habitat confirmation assessments, reporting and mapping. All components leveraged `R`, `SQL` or `Python` programming languages to facilitate workflow tracking, collaboration, transparency and continually improving research. Project workflows utilized local and remote `postgreSQL` databases as well as a "snapshot" of select datasets contained within a local `sqlite` database. A data and script repository to facilitate this reporting is located on [Github](https://github.com/NewGraphEnvironment/fish_passage_elk_2021_reporting).  


```{r eval=F}
source('R/packages.R')
source('R/tables.R')
# or the development version
# devtools::install_github("rstudio/bookdown")
```

## Planning

 
Priorities for site assessment locations were provided by Canadian Wildlife Federation with some additional sites selected by field crews based on background literature, fisheries information, PSCIS, `bcfishpass` [@norris2021smnorrisbcfishpass] outputs and field reconnaissance. `bcfishpass` is an open-source code repository that models aquatic connectivity based on a suite of hard coded (maximum stream slope downstream, PSCIS barrier information, [dam locations](https://github.com/smnorris/bcdams) and user defined parameters ([gradient/width/discharge](https://github.com/smnorris/bcfishpass/tree/main/02_model/parameters_wcrp)).  Details of the general methodology can be found [here](https://github.com/smnorris/bcfishpass) and may be updated over time as the tool evolves.  Once a development environment is properly setup, the software builds a local `postgresql` database through the utilization of other open-source tools such as [`bcdata`](https://github.com/smnorris/bcdata) [@norris2021smnorrisbcdata], [`bcfishobs`](https://github.com/smnorris/bcfishobs) [@norris2021smnorrisbcfishobs], and [`fwapg`](https://github.com/smnorris/fwapg) [@norris2021smnorrisfwapg] to provide connectivity models developed from analysis of the BC Freshwater Atlas, road layers,  fisheries information, Pacific Climate Impacts Consortium stream discharge data, measured and [modeled estimates of channel width](https://www.poissonconsulting.ca/temporary-hidden-link/859859031/channel-width-21b/) and numerous other standardized datasets downloaded directly from the [BC Data Catalogue](https://catalogue.data.gov.bc.ca/) application programming interface using `bcdata`.


 
<br>

### Accessible Habitat 

The `Fish Habitat Model` calculates the average gradient of BC Freshwater Atlas stream network lines at minimum 100m long intervals starting from the downstream end of the streamline segment and working upstream.  The network lines are broken into max gradient categories with new segments created if and when the average slope of the stream line segment exceeds user provided thresholds. For this project, the user provided gradient thresholds used to delineate "potentially accessible habitat" were based on estimated max gradients that salmon (15%) and steelhead (20%) are likely to be capable of ascending. Although not housed within `bcfishpass` at the time of reporting, plans were in place to port over of the latest model version (V2.3.1) of the stream profile analysis tools from `FWAToolsArchive` into `bcfishpass`.  

<br>


Through this initiative, the Provincial Fish Passage Remediation Program and connectivity restoration planning led by Canadian Wildlife Federation and funded by the British Columbia Salmon Restoration and Innovation Fund, `bcfishpass` has been designed to prioritize potential fish passage barriers for assessment or remediation. The software is under continual development and has been designed and constructed by @norris2021smnorrisbcfishpass using of sql and python based shell script libraries to generate a simple model of aquatic habitat connectivity.  The model identifes natural barriers (ex. steep gradients for extended distances) and hydroelectric dams to classifying the accessibility upstream by fish [@norris2021smnorrisbcfishpass]. On potentially accessible streams, scripts identify known barriers (ex. waterfalls >5m high) and additional anthropogenic features which are primarily road/railway stream crossings (i.e. culverts) that are potentially barriers. To prioritize these features for assessment or remediation, scripts report on how much modelled potentially accessible aquatic habitat the barriers may obstruct. The model can be refined with known fish observations upstream of identified barriers and for each crossing location, the area of lake and wetland habitat upstream, species documented upstream/downstream, an estimate of watershed area (on 2nd order and higher streams), mean annual precipitation weighted to upstream watershed area and channel width can be collated using `bcfishpass`, `fwapg` and `bcfishobs`.  This, information, can be used to provides an indication of the potential quantity and quality of habitat potentially gained should fish passage be restored by comparing to user defined thresholds for the aforementioned parameters. A discussion of the methodology to derive channel width is below.
 
 <br>

Gradient, channel size and stream discharge are key determinants of channel morphology and subsequently fish distribution. High value  rearing, overwintering  and spawning habitat preferred by numerous species/life stages of fish are often located within channel types that have relatively low gradients and large channel widths (also quantified by the amount of flow in the stream).  Following delineation of "potentially accessible habitat", the average gradient of each stream segment within habitat classified as below the 15% and 20% thresholds was calculated and summed within species and life stage specific gradient categories.  Average gradient of stream line segments can be calculated from elevations contained in the provincial freshwater atlas streamline dataset. To obtain estimates of channel width upstream of crossing locations, Where available, `bcfishpass` was utilized to pull average channel gradients from Fisheries Information Summary System (FISS) site assessment data [@moeStreamInventorySample] or PSCIS assessment data [@moe2021PSCISAssessments] and associate with stream segment lines. When both FISS and PSCIS values were associated with a particular stream segment, FISS channel width was used.  When multiple FISS sites were associated with a particular stream segment a mean of the average channel widths was taken.  To model channel width for 2nd order and above stream segments without associated FISS or PSCIS sites, first `fwapg` was used to estimate the drainage area upstream of the segment.  Then, rasters from ClimateBC [@wang_etal2012ClimateWNAHighResolution] were sampled for each stream segments and a mean annual precipitation weighted by upstream watershed area was calculated. Mean annual precipitation was then combined with the channel widths and BEC zone information (gathered through a spatial query tied to the bottom of the stream segment) into a dataset (n = 22990) for analysis fo the relationship between these variables.  The details of this analysis and resulting formula used to estimate channel width on stream segments in the Bukley River and Morice River watersheds is included as a technical appendix at `r if(gitbook_on){knitr::asis_output("[here](https://github.com/NewGraphEnvironment/fish_passage_bulkley_2020_reporting/blob/master/docs/channel-width-21.pdf)")} else knitr::asis_output("https://github.com/NewGraphEnvironment/fish_passage_bulkley_2020_reporting/raw/master/docs/channel-width-21.pdf")`. 

<br>



`bcfishpass` and associated tools have been designed to be flexible in analysis, accepting user defined gradient, channel width and stream discharge categories [@moeStreamInventorySample].  Although currently in draft form, and subject to development revisions, gradient and channel width thresholds for habitat with the highest intrinsic value for a number of fish species in the Bulkley River and Morice River watersheds groups have been specified and applied to model habitat upstream of stream crossing locations with the highest intrinsic value (Table \@ref(tab:tab-fish-spawning-rearing)).  Thresholds were derived based on a literature review with references provided in Table \@ref(tab:tab-fish-spawning-rearing-references). 



`r if(gitbook_on){knitr::asis_output("<br>")} else knitr::asis_output("<br><br><br>")`


```{r tab-fish-spawning-rearing}
#`r if(identical(gitbook_on, FALSE)){knitr::asis_output("<br><br><br>")}`
bcfishpass_spawn_rear_model %>% 
  mutate(Species = fishbc::fbc_common_name(species_code), 
         spawn_gradient_max = round(spawn_gradient_max * 100 ,1),
         rear_gradient_max = round(rear_gradient_max * 100 ,1)) %>%
  select(Species, 
         `Spawning Gradient  Max (%)`= spawn_gradient_max,
         `Spawning Width Min (m)` = spawn_channel_width_min,
         # `Spawning Width Max (m)` = spawn_channel_width_max,
         # `Spawning MAD Min (m3/s)` = spawn_mad_min,
         # `Spawning MAD Max (m3/s)` = spawn_mad_max,
         `Rearing Gradient Max (%)` = rear_gradient_max) %>% 
         # `Rearing MAD Min (m3/s)` = rear_mad_min,
         # `Rearing MAD Max (m3/s)` = rear_mad_max,
         # `Rearing Wetland Multiplier` = rear_wetland_multiplier,
         # `Rearing Lake Multiplier` = rear_lake_multiplier) %>% 
  t() %>% 
  as_tibble(rownames = "row_names") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  rename(Variable = Species) %>% 
  my_kable(caption_text = 'Stream gradient and channel width thresholds used to model potentially highest value fish habitat.')

```

<br>

```{r tab-fish-spawning-rearing-references}
# bcfishpass_spawn_rear_model_references <- readr::read_csv(file = 'data/width_modelling/model_spawning_rearing_habitat.csv')
bcfishpass_spawn_rear_model_references <- readr::read_csv(file = 'data/width_modelling/model_spawning_rearing_habitat_ref.csv') %>% 
  select(species_code, contains('ref'), -contains(c('multiplier','mad')))
  
bcfishpass_spawn_rear_model_references %>% 
  mutate(Species = fishbc::fbc_common_name(species_code)) %>% 
  select(Species, 
         `Spawning Gradient  Max (%)`= spawn_gradient_max_ref,
         `Spawning Width Min (m)` = spawn_channel_width_min_ref,
         # `Spawning Width Max (m)` = spawn_channel_width_max_ref,
         # `Spawning MAD Min (m3/s)` = spawn_mad_min,
         # `Spawning MAD Max (m3/s)` = spawn_mad_max,
         `Rearing Gradient Max (%)` = rear_gradient_max_ref) %>% 
         # `Rearing Wetland Multiplier` = rear_wetland_multiplier,
         # `Rearing Lake Multiplier` = rear_lake_multiplier) %>% 
         # `Rearing MAD Min (m3/s)` = rear_mad_min,
         # `Rearing MAD Max (m3/s)` = rear_mad_max) %>% 
  t() %>% 
  as_tibble(rownames = "row_names") %>% 
  janitor::row_to_names(row_number = 1) %>% 
  rename(Variable = Species) %>% 
  my_kable(caption_text = 'References for stream gradient and channel width thresholds used to model potentially highest value fish habitat. Preliminary and subject to revisions.')
```









```{r tablethreshaverage, eval= F}
#to quantify upstream habitat potentially available for salmonids and facilitate stream line symbology based on stream morphology.
# while high gradient sections typically  present  upstream  migration  barriers  and  less  available  habitat.  Additionally, the size of the stream (indicated by channel width) is an important determinant for habitat suitability for different species as well as specific life stages of those species. 

# `bcfishpass` was used to categorize and sum potentially accessible stream segments in the study area watersheds within gradient and width categories for each stream segment. 
# (0 - 3%, 3 - 5%, 5 - 8%, 8 - 15%, 15 - 20%) with these outputs further amalgamated to summarize and symbolize potential upstream habitat in three categories: riffle/cascade (0 - 5%), step-pool (5 - 15%) and step-pool very steep (15-20%) (Table \@ref(tab:tablethreshaverage)).  


#threshold and average gradient table
table_thresh_average <- tibble::tibble(`Gradient` = c('0 - 5%', '5 - 15%', '15 - 20%', '>20%'),
                                       `Channel Type` = c('Riffle and cascade pool', 'Step pool', 'Step pool - very steep', 'Non fish habitat'))

table_thresh_average %>% 
    my_kable(caption_text = 'Stream gradient categories (threshold and average) and associated channel type.')

```


<br>


### PSCIS and Modelled Stream Crossing Review

To prepare for Phase 1 and 2 assessments in the study area, past fish passage assessment reports for the Bulkley River and Morice River watershed groups were first reviewed to identify crossing structures not yet assessed or previously ranked as priorities for rehabilitation [@casselman_stanley2010BulkleyFulton; @irvine2018AnalysisPriority; @mccarthy_fernando20152015Inventory; @smith2018AssessingBarriers; @wilson_rabnett2007FishPassage].  To determine which of those crossings had not yet been assessed with Phase 1 and Phase 2 assessments we cross-referenced these reports with the PSCIS database, available background info and viewed sites within the output of the `Fish Habitat Model` and `bcfishpass`.  Outputs for modelled and PSCIS crossings (barriers and potential barriers) that met the following criteria underwent a detailed review to facilitate prioritization for Phase1 - Fish Passage Assessments and Phase 2 - Habitat Confirmations.

 * Confirmed fish presence upstream of the structure.
 * Stream width documented as > 2.0m in PSCIS.
 * Linear lengths of modelled upstream habitat <8% gradient for ≥1km. 
 * Crossings located on streams classified as 3rd order or higher.
 * Crossings located on streams with >5 ha of modeled wetland and/or lake habitat upstream.
 * Habitat value rated as "medium" or “high” in PSCIS.  
 

<br>


## Fish Passage Assessments

In the field, crossings prioritized for follow-up were first assessed for fish passage following the procedures outlined in “Field Assessment for Determining Fish Passage Status of Closed Bottomed Structures” [@fish_passage_assessments]. Crossings surveyed included closed bottom structures (CBS), open bottom structures (OBS) and crossings considered “other” (i.e. fords).  Photos were taken at surveyed crossings and when possible included images of the road, crossing inlet, crossing outlet, crossing barrel, channel downstream and channel upstream of the crossing and any other relevant features.  The following information was recorded for all surveyed crossings: date of inspection, crossing reference, crew member initials, Universal Transverse Mercator (UTM) coordinates, stream name, road name and kilometer, road tenure information, crossing type, crossing subtype, culvert diameter or span for OBS, culvert length or width for OBS.  A more detailed “full assessment” was completed for all closed bottom structures and included the following parameters: presence/absence of continuous culvert embedment (yes/no), average depth of embedment, whether or not the culvert bed resembled the native stream bed, presence of and percentage backwatering, fill depth, outlet drop, outlet pool depth, inlet drop, culvert slope, average downstream channel width, stream slope, presence/absence of beaver activity, presence/absence of fish at time of survey, type of valley fill, and a habitat value rating.  Habitat value ratings were based on channel morphology, flow characteristics (perennial, intermittent, ephemeral), fish migration patterns, the presence/absence of deep pools, un-embedded boulders, substrate, woody debris, undercut banks, aquatic vegetation and overhanging riparian vegetation (Table \@ref(tab:tab-hab-value)).  For crossings determined to be potential barriers or barriers based on the data (see [Barrier Scoring]), a culvert fix and recommended diameter/span was proposed.  

<br>


 
```{r tab-hab-value}
tab_habvalue %>% 
  knitr::kable(caption = 'Habitat value criteria (Fish Passage Technical Working Group, 2011).', booktabs = T) %>% 
    kableExtra::column_spec(column = 1, width_min = '1.5in') %>% 
    kableExtra::kable_styling(c("condensed"), full_width = T, font_size = font_set) 
  
```
 
`r if(gitbook_on){knitr::asis_output("<br>")} else knitr::asis_output("\\pagebreak")`

### Barrier Scoring

Fish passage potential was determined for each stream crossing identified as a closed bottom structure as per @fish_passage_assessments.  The combined scores from five criteria: depth and degree to which the structure is embedded, outlet drop, stream width ratio, culvert slope, and culvert length were used to screen whether each culvert was a likely barrier to some fish species and life stages (Table \@ref(tab:tab-barrier-scoring), Table \@ref(tab:tab-barrier-result). These criteria were developed based on data obtained from various studies and reflect an estimation for the passage of a juvenile salmon or small resident rainbow trout [@clarkinNationalInventoryAssessment2005 ;@bellFisheriesHandbookEngineering1991; @thompsonAssessingFishPassage2013].  

<br>

```{r tab-barrier-scoring, eval=T}
tab <- as_tibble(t(tab_barrier_scoring)) %>% 
  mutate(V4 = names(tab_barrier_scoring)) %>% 
  select(V4, everything()) %>% 
  janitor::row_to_names(1) %>%  ##turn the table sideways
  mutate(Risk = case_when(Risk == 'Value' ~ '  Value',
                          T ~ Risk))

tab %>% 
  my_kable(caption_text = 'Fish Barrier Risk Assessment (MoE 2011).')

```

<br>


```{r tab-barrier-result}
tab_barrier_result %>% 
  my_kable(caption_text = 'Fish Barrier Scoring Results (MoE 2011).') 

```

<br>

### Cost Benefit Analysis

A cost benefit analysis was conducted for each crossing determined to be a barrier based on an estimate of cost associated with remediation or replacement of the crossing with a structure that facilitates fish passage and the amount of potential habitat that would be made available by remediating fish passage at the site (habitat gain index).  

<br>

#### Habitat Gain Index

The habitat gain index is the quantity of modelled habitat upstream of the subject crossing and represents an estimate of habitat gained with remediation of fish passage at the crossing.  For this project, a gradient threshold between accessible and non-accessible habitat was set at 20% (for a minimimum length of 100m) intended to represent the maximum gradient of which the strongest swimmers of anadromous species (steelhead) are likely to be able to migrate upstream.  

<br>

For reporting of Phase 1 - fish passage assessments within the body of this report (Table \@ref(tab:tab-barrier-scoring)), a "total" value of habitat <20% output from `bcfishpass` was used to estimate the amount of habitat upstream of each crossing less than 20% gradient before a falls of height >5m - as recorded in @ProvincialObstaclesFish or documented in other `bcfishpass` online documentation. To generate areas of habitat upstream, the estimated linear length was multiplied by the downstream channel width measured as part of the fish passage assessment protocol.  Although these estimates are not generally conservative, have low accuracy and do not account for upstream stream crossing structures they do allow a rough screening of the best candidates for follow up with more detailed Phase 2 assessments. 

<br>

For Phase 2 - habitat confirmation sites, conservative estimates of the linear quantity of habitat to be potentially gained by fish passage restoration, mainstem and large tributary streams (>1st order streams) segments upstream of each crossing that were <20%, below natural barriers and downstream of documented culvert barriers were measured by hand with the measure tool within QGIS [@QGIS_software]. To generate estimates of the area of habitat upstream of these sites, the length of habitat was multiplied by the upstream average channel width that was measured in the field.

<br>


Potential options to remediate fish passage were selected from @fish_passage_assessments and included:  

 + Removal (RM) - Complete removal of the structure and deactivation of the road. 
 + Open Bottom Structure (OBS) - Replacement of the culvert with a bridge or other open bottom structure.  For this project we considered bridges as the only viable option for OBS type based on consultation with FLNR road crossing engineering experts.  It should be noted however, that box culverts could be considered a viable and economical option as they have been observed as successfully facilitating fish passage on the west coast of the province (Betty Rebellato, Canadian Wildlife Federation - Project Biologist).
 + Streambed Simulation (SS) - Replacement of the structure with a streambed simulation design culvert.  Often achieved by embedding the culvert by 40% or more. Based on consultation with FLNR engineering experts, we considered crossings on streams with a channel width of <2m and a stream gradient of <8% as candidates for replacement with streambed simulations.
 + Additional Substrate Material (EM) - Add additional substrate to the culvert and/or downstream weir to embed culvert and reduce overall velocity/turbulence.  This option was considered only when outlet drop = 0, culvert slope <1.0% and stream width ratio < 1.0.
 + Backwater (BW) - Backwatering of the structure to reduce velocity and turbulence. This option was considered only when outlet drop < 0.3m, culvert slope <2.0%, stream width ratio < 1.2 and stream profiling indicates it would be effective..
 
 <br>

Cost estimates for structure replacement with bridges and embedded culverts were generated based on the channel width, slope of the culvert, depth of fill, road class and road surface type. Road details were sourced from @flnrordForestTenureRoad2020 and @flnrordDigitalRoadAtlas2020 through `bcfishpass`. Interviews with Phil MacDonald, Engineering Specialist FLNR - Kootenay, Steve Page, Area Engineer - FLNR - Northern Engineering Group and Matt Hawkins - MoTi - Design Supervisor for Highway Design and Survey - Nelson were utilized to helped refine estimates.  

<br>

Base costs for installation of bridges on forest service roads and permit roads with surfaces specified in provincial GIS road layers as rough and loose was estimated at \$12500/linear m and assumed that the road could be closed during construction and a minimum bridge span of 10m. For streams with channel widths <2m, embedded culverts were reported as an effective solution with total installation costs estimated at $25k/crossing (pers. comm. Phil MacDonald, Steve Page). For larger streams (>6m), span width increased proportionally to the size of the stream (ex. for an 8m wide stream a 12m wide span was prescribed).  For crossings with large amounts of fill (>3m), the replacement bridge span was increased by an additional 3m for each 1m of fill >3m to account for cutslopes to the stream at a 1.5:1 ratio. To account for road type, a multiplier table was also generated to estimate incremental cost increases with costs estimated for structure replacement on paved surfaces, railways and arterial/highways costing up to 20 times more than forest service roads due to expenses associate with design/engineering requirements, traffic control and paving.  The cost multiplier table (Table \@ref(tab:tab-cost-mult)) should be considered very approximate with refinement recommended for future projects.  

`r if(gitbook_on){knitr::asis_output("<br>")} else knitr::asis_output("\\pagebreak")`

```{r tab-cost-mult}
# print_tab_cost_mult(caption_text = 'Cost multiplier table based on road class and surface type.')
tab_cost_rd_mult_report %>%
  my_kable(caption_text = 'Cost multiplier table based on road class and surface type.')
```

<br>

## Habitat Confirmation Assessments

Following fish passage assessments, habitat confirmations were completed in accordance with procedures outlined in the document “A Checklist for Fish Habitat Confirmation Prior to the Rehabilitation of a Stream Crossing” [@confirmation_checklist_2011]. The main objective of the field surveys was to document upstream habitat quantity and quality and to determine if any other obstructions exist above or below the crossing.  Habitat value was assessed based on channel morphology, flow characteristics (perennial, intermittent, ephemeral), the presence/absence of deep  pools, un-embedded  boulders, substrate, woody debris, undercut banks, aquatic vegetation and overhanging riparian vegetation. Criteria used to rank habitat value was based on guidelines in @confirmation_checklist_2011 (Table \@ref(tab:tab-hab-value)). 

<br>

During habitat confirmations, to standardize data collected and facilitate submission of the data to provincial databases, information was collated on ["Site Cards"](https://www2.gov.bc.ca/gov/content/environment/natural-resource-stewardship/laws-policies-standards-guidance/inventory-standards/aquatic-ecosystems). Habitat characteristics recorded included channel widths, wetted widths, residual pool depths, gradients, bankfull depths, stage, temperature, conductivity, pH, cover by type, substrate and channel morphology (among others). When possible, the crew surveyed downstream of the crossing to the point where fish presence had been previously confirmed and upstream to a minimum distance of 600m. Any  potential  obstacles  to  fish  passage  were  inventoried  with  photos, physical  descriptions  and  locations  recorded on site cards.  Surveyed routes were recorded with time-signatures on handheld GPS units.   

<br>

Fish sampling was conducted a subset of sites when biological data was considered to add significant value to the physical habitat assessment information. When possible, electrofishing was utilized within discrete site units both upstream and downstream of the subject crossing with electrofisher settings, water quality parameters (i.e. conductivity, temperature and ph), start location, length of site and wetted widths (average of a minimum of three) recorded.  For each fish captured, fork length and species was recorded, with results included within the fish data submission spreadsheet.  Fish information and habitat data will be submitted to the province under scientific fish collection permit CB20-611971.

<br>

\pagebreak

## Reporting

Reporting was generated with `bookdown` [@bookdown2016] from `Rmarkdown` [@R-rmarkdown] with primarily `R` [@rcoreteam2020languageenvironment] and `SQL` scripts. In addition to numerous spatial layers sourced through the BC Data Catalogue then stored and queried in a local `postgresql` database [data inputs](https://github.com/NewGraphEnvironment/fish_passage_bulkley_2020_reporting/tree/master/data) for this project include: 

 + Populated [Fish Data Submission Spreadsheet Template - V 2.0, January 20, 2020 ](https://www2.gov.bc.ca/gov/content/environment/plants-animals-ecosystems/fish/fish-and-fish-habitat-data-information/fish-data-submission/submit-fish-data#submitfish) 

 + Populated [pscis_assessment_template_v24.xls](https://www2.gov.bc.ca/gov/content/environment/plants-animals-ecosystems/fish/aquatic-habitat-management/fish-passage/fish-passage-technical/assessment-projects)
 
 
 + [`Fish Habitat Model`/`bcfishpass`](https://github.com/NewGraphEnvironment/fish_passage_bulkley_2020_reporting/blob/master/data/bcfishpass-phase2.csv) outputs.


 + [Custom CSV file](https://github.com/NewGraphEnvironment/fish_passage_bulkley_2020_reporting/raw/master/data/habitat_confirmations_priorities.csv) detailing Phase 2 site:
     - priority level for proceeding to design for replacement
     - length of survey upstream and downstream
     - a conservative estimate of the linear length of mainstem habitat potentially available upstream of the crossing 
     - fish species confirmed as present upstream of the crossing

 + [GPS tracks](https://github.com/NewGraphEnvironment/fish_passage_bulkley_2020_reporting/tree/master/data/habitat_confirmation_tracks.gpx) from field surveys.  

 + [Photos](https://github.com/NewGraphEnvironment/fish_passage_bulkley_2020_reporting/tree/master/data/photos) and [photo metadata](https://github.com/NewGraphEnvironment/fish_passage_bulkley_2020_reporting/tree/master/data/photo_metadata.csv)


## Mapping

Mapping was completed by Hillcrest Geographics.  `pdf` maps were generated using `QGIS` with data supplied via a `postgreSQL` database.  A QGIS layer file defining and symbolizing all layers required for general fish passage mapping was developed and at the time of reporting was kept under version control within `bcfishpass`.
 

 





